### 2016-11-11 Paper_v1.tex
Nov11_3

\subsection*{Estimation of an ensemble of cell-free protein synthesis models.}
We used the hybrid cell-free modeling framework of Wayman et al. to simulate the production of a model protein [REFHERE].
The cell-free \textit{E. coli} metabolic model was constructed by removing the growth-associated processes from the model of Palsson and coworkers \cite{2000_edwards_palsson_PNAS}, and by adding reactions for the synthesis of chloramphenicol acetyltransferase (CAT), a model protein for which we have a comprehensive training dataset \cite{2005_calhoun_BiotechnologyProgress}.
Thus, the model described core central carbon metabolism (glycolysis, pentose phosphate, Enter-Doudoroff, TCA cycle),
as well as the synthesis of energy species, amino acids biosynthesis and degradation, and biosynthesis of the CAT protein.
An ensemble of model parameters was estimated from dynamic measurements of glucose, CAT, organic acids (pyruvate, lactate, acetate, succinate, malate), energy species (A(x)P, G(x)P, C(x)P, U(x)P), and 18 of the 20 proteinogenic amino acids. We generated an ensemble of N = 18,000 parameter sets by minimizing the error between the training dataset and the metabolite concentrations predicted by the model. We defined the set with the lowest error value as the best-fit parameter set. [STATISTICS ON PARAMETERS].

The ensemble of models captured the time evolution of cell-free CAT biosynthesis (Fig.~\ref{fig:CarbonBoth} - \ref{fig:Amino}).
Glucose was exhausted with 3 hr [FILL ME IN].
The ensemble also captured the energy species dynamics, particularly the overall energy total (Fig.~\ref{fig:CarbonBoth}, top) and the totals by base .
The ensemble and the best-fit set also predict some of the amino acid measurements, while failing to predict others (Fig.~\ref{fig:Amino}).
the central carbon metabolism, including glucose uptake, CAT production, and the dynamics of the organic acid intermediates .
Allosteric control is important to the dynamics of the organic acid intermediates, as without it several of the measurements are not captured by the ensemble or the best-fit set (Fig.~\ref{fig:CarbonBoth}, bottom).
This is likely due to a structural deficiency in the model; in some cases, the consumption of an amino acid through CAT synthesis is not enough to explain the decrease shown in the data, and there are no other reactions that consume it. Thus, a more comprehensive biological description is needed to fully explain amino acid dynamics.

@@

\subsection*{Sensitivity analysis}
We determined the reactions most important to protein production by computing the local sensitivity of CAT concentration to each rate constant in the network.
Each sensitivity index was formulated as:
\begin{equation}\label{eqn:local-sensitivity}
\begin{split}
    S_{ij}=norm(CAT(p_i,p_j,t)-CAT(\alpha*p_i,\alpha*p_j,t))\qquad{i,j=1,2,\hdots\mathcal{P}}\\
\end{split}
\end{equation}
where $S_{ij}$ denotes the sensitivity of CAT production to the $i$th and $j$th parameters, $CAT(p_i,p_j,t)$ denotes CAT concentration as a function of time and the $i$th and $j$th parameters, $\alpha$ denotes the perturbation factor, equal to 1.01, and $\mathcal{P}$ denotes the number of rate constants, equal to 163.
In calculating the pairwise sensitivities, each parameter was perturbed by 1\%; first-order sensitivities ($i$ = $j$) were subject to two 1\% perturbations, equivalent to a perturbation of 2.01\%.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth,trim=0 28 615 40,clip]{./Figures/Sensitivity.pdf}
\caption{Normalized first-order and pairwise sensitivities of CAT production to rate constants.}
\label{fig:Sensitivity}
\end{figure}

>>>>

\subsection*{Estimation of an ensemble of cell-free protein synthesis models.}
We used the hybrid cell-free modeling framework of Wayman et al. to simulate the production of a model protein [REFHERE].
The cell-free \textit{E. coli} metabolic model was constructed by removing the growth-associated processes from the model of Palsson and coworkers \cite{2000_edwards_palsson_PNAS}, and by adding reactions for the synthesis of chloramphenicol acetyltransferase (CAT), a model protein for which we have a comprehensive training dataset \cite{2005_calhoun_BiotechnologyProgress}.
Thus, the model described core central carbon metabolism (glycolysis, pentose phosphate, Enter-Doudoroff, TCA cycle),
as well as the synthesis of energy species, amino acids biosynthesis and degradation, and biosynthesis of the CAT protein.
An ensemble of model parameters was estimated from dynamic measurements of glucose, CAT, organic acids (pyruvate, lactate, acetate, succinate, malate), energy species (A(x)P, G(x)P, C(x)P, U(x)P), and 18 of the 20 proteinogenic amino acids.
We generated an ensemble of N = 100 diverse parameter sets by minimizing the error between the training dataset and the metabolite concentrations predicted by the model, and choosing sets so as to minimize the Pearson correlation coefficient between sets in the ensemble.

The ensemble of models captured the time evolution of cell-free CAT biosynthesis (Fig.~\ref{fig:CarbonBoth} - \ref{fig:Amino}).
The ensemble and the best-fit set also captured the energy species dynamics, particularly the overall energy total (Fig.~\ref{fig:CarbonBoth}, top) and the totals by base, as well as the amino acid measurements (Fig.~\ref{fig:Amino}).
Allosteric control is important to the dynamics of the organic acid intermediates, as without it several of the measurements are not captured by the ensemble or the best-fit set (Fig.~\ref{fig:CarbonBoth}, bottom).

@@

\subsection*{Local sensitivity analysis}
We determined the reactions most important to protein production by computing the local sensitivity of CAT concentration to each rate constant in the network.
Each sensitivity index was formulated as:
\begin{equation}\label{eqn:local-sensitivity}
\begin{split}
    S_{ij}=norm(CAT(p_i,p_j,t)-CAT(\alpha*p_i,\alpha*p_j,t))\qquad{i,j=1,2,\hdots\mathcal{P}}\\
\end{split}
\end{equation}
where $S_{ij}$ denotes the sensitivity of CAT production to the $i$th and $j$th parameters, $CAT(p_i,p_j,t)$ denotes CAT concentration as a function of time and the $i$th and $j$th parameters, $\alpha$ denotes the perturbation factor, equal to 1.01, and $\mathcal{P}$ denotes the number of rate constants, equal to 163.
In calculating the pairwise sensitivities, each parameter was perturbed by 1\%; first-order sensitivities ($i$ = $j$) were subject to two 1\% perturbations, equivalent to a perturbation of 2.01\%.
The array of sensitivity indices was organized into a clustergram (Fig.~\ref{fig:Sensitivity}, top) with both axes corresponding to the 163 rate constants rearranged from most sensitive (top, right) to least sensitive (bottom, left).

Likewise, we determined the reactions most important to the system as a whole by computing the local sensitivity of the species for which data exist to each rate constant in the network.
Each sensitivity index was formulated as:
\begin{equation}\label{eqn:local-sensitivity}
\begin{split}
    S_{ij}=norm(X(p_i,p_j,t)-X(\alpha*p_i,\alpha*p_j,t))\qquad{i,j=1,2,\hdots\mathcal{P}}\\
\end{split}
\end{equation}
where $S_{ij}$ denotes the sensitivity of the system state to the $i$th and $j$th parameters, and $X(p_i,p_j,t)$ denotes the system state, an array consisting of the concentration of every species for which data exist as a function of time and the $i$th and $j$th parameters.
% $\alpha$ denotes the perturbation factor, equal to 1.01, and $\mathcal{P}$ denotes the number of rate constants, equal to 163.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth,trim=0 28 615 40,clip]{./Figures/Sensitivity.pdf}
\caption{Normalized first-order and pairwise sensitivities of CAT production (top) and system state (bottom) to rate constants.}
\label{fig:Sensitivity}
\end{figure}
